---
title: "SIPS Multiverse"
subtitle: "2. Descriptive Specification Curve"
author: "Constantin Yves Plessen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
knit: (function(inputFile, encoding) {
      rmarkdown::render(inputFile, 
      encoding = encoding, 
      output_dir = "html") })
output:
  html_document:
    code_folding: show
    highlight: pygment
    theme: cerulean
    toc: yes
    toc_depth: 2
    toc_float: yes
---

```{r setup, include=FALSE, out.width="100%"}
knitr::opts_chunk$set(echo = TRUE)
library(cowplot)
library(ggrepel)
library(tidyverse) # data cleaning
```

# Descriptive Specification Plot

```{r}
specifications_full <- read.csv2("data/specifications_full.csv")[, -1]
data_sim <- read.csv2("data/data_sim.csv")[, -1]
```


### Preparation

<br>

### Which factors


### Which factors
```{r}
wf1   <- c(unique(data_sim$wf1), "all wf1")
outliers <- c("not_outlier", "all effects")
rob <- c("exclude_worst", "include_best", "all rob")
```


### How Factors
```{r}
dependency <- c("ignore", "aggregate", "modeled")

ma_method     <- c("3-level", "rve",       # modeled
                   "reml", "fe",           # aggregate | ignore
                   "puniform", "pet-peese" # aggregate only
                   )

number_which_how_factors <- 5
```

#### Rank each summary effect size by magnitude

```{r }
x_rank <- rank(specifications_full$mean, 
               ties.method = "random")
```


#### Create all factors
```{r }
yvar <- rep(factor(rev(c(
  wf1,
  outliers,
  rob,
  dependency,
  ma_method )), levels = rev(c(
    wf1,
    outliers,
    rob,
    dependency,
    ma_method ))), 
  times = nrow(specifications_full))

unique(yvar)
```


#### Check lengths of yvar and specifications
```{r }
length(yvar)

length(levels(yvar)) * nrow(specifications_full)
```


```{r }
xvar <- rep(x_rank, 
            each = length(levels(yvar)))
spec <- NULL
```

<br>

#### Determine which specifications are observed and which are not

```{r }
for(i in 1:nrow(specifications_full)) {
  id <- as.numeric(levels(yvar) %in% 
                     as.character(unlist(
                       specifications_full[i, 1:number_which_how_factors])))  
  spec <- c(spec, id)
}
```


```{r }
plotdata <- data.frame(xvar, 
                       yvar, 
                       spec)

head(plotdata)
```

<br>

#### Define y axis labels

Need to be in order of:
```{r}
rev(unique(yvar))
```

```{r }
ylabels <- rev(c(
"Group: Non-adults",  
"Group: Adults",
"Group: All groups", 

"Outlier: Removed",
"Outlier: Not Removed", 

"ROB: Worst Excluded",  
"ROB: Only Best",
"ROB: All rob", 

"Dependency: Ignore", 
"Dependency: Aggregate",
"Dependency: Modeled",

"Method: 3-Level",
"Method: RVE",
"Method: REML",
"Method: FE",
"Method: p-uniform",
"Method: PET-PEESE"
 ))

ylabels
```


<br>

```{r }
plotdata$k <- rep(specifications_full$k, 
                  each = length(levels(yvar)))  

plotdata$fill <- as.factor(plotdata$k * plotdata$spec)
```

```{r}
write.csv(plotdata, "data/plotdata.csv")
```

<br>

## Colors based on relative size (*k*)
```{r}
fill_quantiles <- quantile(plotdata$k, c(.10, .20, .30, .40, .50, .60, .70, .80, .90))
```

```{r}
plotdata_rel <- plotdata %>% 
  mutate(fill_manual = case_when(
    spec == 0 ~ 0,           # white
    k  != 0               & k <= fill_quantiles[1] ~ 1,     # light yellow
    k > fill_quantiles[1] & k <= fill_quantiles[2] ~ 2,
    k > fill_quantiles[2] & k <= fill_quantiles[3] ~ 3,
    k > fill_quantiles[3] & k <= fill_quantiles[4] ~ 4,
    k > fill_quantiles[4] & k <= fill_quantiles[5] ~ 5,
    k > fill_quantiles[5] & k <= fill_quantiles[6] ~ 6,
    k > fill_quantiles[6] & k <= fill_quantiles[7] ~ 7,
    k > fill_quantiles[7] & k <= fill_quantiles[8] ~ 8,
    k > fill_quantiles[8] & k <= fill_quantiles[9]  ~ 9,
    k > fill_quantiles[9]  ~ 10),   # dark red
    fill_manual = as.factor(fill_manual))
```


```{r }
cols <- RColorBrewer::brewer.pal(min(11, length(levels(plotdata_rel$fill_manual)) - 1), "Spectral")

plot(NULL, xlim=c(0,length(cols)), ylim=c(0,1), 
    xlab=cols, ylab="", xaxt="n", yaxt="n")

rect(0:(length(cols)-1), 0, 1:length(cols), 1, col=cols)
```

##### Reversed color palette
```{r}
cols_rev <- cols[floor(seq(from = length(cols), to = 0, # change from to to reverse color coding!
              length.out = length(levels(plotdata_rel$fill_manual))))] # change - 1 for direction change
plot(NULL, xlim=c(0,length(cols_rev)), ylim=c(0,1), 
    xlab="", ylab="", xaxt="n", yaxt="n")
rect(0:(length(cols_rev)-1), 0, 1:length(cols_rev), 1, col=cols_rev)
```

<br>

##### Specification tile plot 

We will need the length of each factor when we plot the spec curve:

```{r}
length_of_each_factor <- c(
  length(ma_method),
  length(dependency) + length(ma_method),
  length(rob) + length(dependency) + length(ma_method),
  length(rob) + length(dependency) + length(ma_method) + length(outliers),
  length(outliers) + length(wf1) + length(rob) + length(dependency) + length(ma_method))
```

### Empty factor pane

```{r}
factor_pane <- ggplot(data = plotdata_rel, 
       aes(x = xvar, 
           y = as.factor(yvar), 
           fill = "white")) +
  geom_raster(aes(alpha = 0)) + 
  geom_hline(yintercept = length_of_each_factor + 0.5) +  # Change lines here here
  scale_x_continuous(position = "bottom") +
  scale_y_discrete(labels = ylabels) +
  scale_fill_manual(
    values = c("white", cols_rev)) +
  labs(x = "Specification number", 
       y = "Which/How factors") +
  coord_cartesian(
    expand = F, xlim = c(0.5, nrow(specifications_full) + 0.5)) +
  theme_bw() + 
  theme(legend.position = "none",
        axis.text.y = element_text(colour = "black", size = 8),
        axis.text.x = element_text(colour = "black"),
                axis.ticks = element_line(colour = "black"),
        plot.margin = margin(t = 5.5, 
                             r = 5.5, 
                             b = 5.5, 
                             l = 5.5, 
                             unit = "pt"))

ggsave("figures/factor_pane.png", 
       factor_pane, 
       width = 10, 
       height = 7.5,
       dpi = "retina"
       )
```

```{r}
p1_rel <- ggplot(data = plotdata_rel, 
       aes(x = xvar, 
           y = as.factor(yvar), 
           fill = fill_manual)) +
  geom_raster() + 
  geom_hline(yintercept = length_of_each_factor + 0.5) +  # Change lines here here
  scale_x_continuous(position = "bottom") +
  scale_y_discrete(labels = ylabels) +
  scale_fill_manual(
    values = c("white", cols_rev)) +
  labs(x = "Specification number", 
       y = "Which/How factors") +
  coord_cartesian(
    expand = F, xlim = c(0.5, nrow(specifications_full) + 0.5)) +
  theme_bw() + 
  theme(legend.position = "none",
        axis.text.y = element_text(colour = "black", size = 8),
        axis.text.x = element_text(colour = "black"),
                axis.ticks = element_line(colour = "black"),
        plot.margin = margin(t = 5.5, 
                             r = 5.5, 
                             b = 5.5, 
                             l = 5.5, 
                             unit = "pt"))

p1_rel
```

```{r}
fill_quantiles <- quantile(specifications_full$k, c(.10, .20, .30, .40, .50, .60, .70, .80, .90))

specifications_full_rel <- specifications_full %>% 
  mutate(fill_manual = case_when(
    k  != 0               & k <= fill_quantiles[1] ~ 1,     # light yellow
    k > fill_quantiles[1] & k <= fill_quantiles[2] ~ 2,
    k > fill_quantiles[2] & k <= fill_quantiles[3] ~ 3,
    k > fill_quantiles[3] & k <= fill_quantiles[4] ~ 4,
    k > fill_quantiles[4] & k <= fill_quantiles[5] ~ 5,
    k > fill_quantiles[5] & k <= fill_quantiles[6] ~ 6,
    k > fill_quantiles[6] & k <= fill_quantiles[7] ~ 7,
    k > fill_quantiles[7] & k <= fill_quantiles[8] ~ 8,
    k > fill_quantiles[8] & k <= fill_quantiles[9]  ~ 9,
    k > fill_quantiles[9]  ~ 10),   # dark red
  fill_manual = as.factor(fill_manual))

specifications_full_rel$xvar <- x_rank

yrng <- range(c(0, specifications_full$lb, specifications_full$ub))

#ylimit <- round(c(yrng[1] - diff(yrng)*0.1, yrng[2] + diff(yrng)*0.1), 1)
ylimit <- c(-.5, 2)

#y_breaks_forest <- round(seq(from = round(ylimit[1], 1), to = round(ylimit[2], 1), by = 0.5), 2)
y_breaks_forest <- seq(-.25, 2, 0.25)

y_labels_forest <- format(y_breaks_forest, nsmall = 2)
y_breaks_forest <- c(ylimit[1], y_breaks_forest)
y_labels_forest <- c(ylabels[which.max(nchar(ylabels))], y_labels_forest)
```

### Select specific Meta-Analyses to show in Plot
```{r}
set.seed(42)
ma_for_forest_1 <- specifications_full_rel %>% 
  filter(full_set == 1 & ma_method == "3-level") %>% mutate(label = "All Specifications: 3-LVL")

ma_for_forest_2 <- specifications_full_rel %>% 
  arrange(mean) %>% 
  slice(which.min(abs(mean - 0))) %>% 
  mutate(label = "Meta-Analysis A")

ma_for_forest_3  <- specifications_full_rel %>% 
  arrange(mean) %>% 
  slice(which.min(abs(lb - 0))) %>% 
  mutate(label = "Meta-Analysis B")

ma_for_forest_4  <- specifications_full_rel %>% 
  arrange(mean) %>% 
  slice(which.min(abs(mean - 2))) %>% 
  mutate(label = "Meta-Analysis C")

ma_for_forest <- bind_rows(ma_for_forest_1,
                           ma_for_forest_2,
                           ma_for_forest_3,
                           ma_for_forest_4)
```

## Empty which/how factor pane
```{r}
effect_pane <- specifications_full_rel %>% 
 # filter(!ma_method == "rve") %>% 
  ggplot(aes(x = xvar, 
             y = mean))+ 
  geom_errorbar(aes(ymin = lb, 
                    ymax = ub,
                    col = as.factor(fill_manual)), 
                width = 0, 
                size = 1,
                alpha = 0) +
  geom_line(col = "black", size = 0.25, alpha = 0) +
  geom_hline(yintercept = 0, color = "blue", linetype = "solid", size = 2) +
  #geom_hline(yintercept = .24, linetype = "dotted", size = 0.7, color = "red") +
  scale_x_continuous(name = "") +
  scale_y_continuous(name = expression(paste("Summary effect (", italic("g"),")")),
                     breaks = y_breaks_forest, labels = y_labels_forest) + 
  scale_color_manual(values = cols_rev) +
  coord_cartesian(ylim = ylimit, xlim = c(0.5, nrow(specifications_full) + 0.5), expand = FALSE) +
  ggtitle("Multiverse of All Reasonable Meta-Analyses") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(colour = c("white", rep("black", times = length(y_labels_forest) - 1))),
        axis.ticks.y = element_line(colour = c("white", rep("black", times = length(y_breaks_forest) - 1))),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(),
        panel.grid.minor.y = element_blank(),
        plot.margin = margin(t = 5.5, r = 5.5, b = -15, l = 5.5, unit = "pt"))
  #geom_label_repel(data = ma_for_forest, 
  #                 aes(label = label),
  #                 seed = 42,
  #                 box.padding   = 0.35, 
  #                 point.padding = 0.5,
  #                 direction = "both",
  #                 lineheight = 2,
  #                 nudge_y = .5,
  #                 segment.color = 'black',
  #                 segment.angle = 90,
  #                 segment.size = .5)

ggsave("figures/effect_pane.png", 
       effect_pane, 
       width = 10, 
       height = 7.5,
       dpi = "retina"
       )
```

```{r}
p2_rel <- specifications_full_rel %>% 
 # filter(!ma_method == "rve") %>% 
  ggplot(aes(x = xvar, 
             y = mean))+ 
  geom_errorbar(aes(ymin = lb, 
                    ymax = ub,
                    col = as.factor(fill_manual)), 
                width = 0, 
                size = 1,
                alpha = .9) +
  geom_line(col = "black", size = 0.25) +
  geom_hline(yintercept = 0, color = "blue", linetype = "solid", size = 2) +
  #geom_hline(yintercept = .24, linetype = "dotted", size = 0.7, color = "red") +
  scale_x_continuous(name = "") +
  scale_y_continuous(name = expression(paste("Summary effect (", italic("g"),")")),
                     breaks = y_breaks_forest, labels = y_labels_forest) + 
  scale_color_manual(values = cols_rev) +
  coord_cartesian(ylim = ylimit, xlim = c(0.5, nrow(specifications_full) + 0.5), expand = FALSE) +
  ggtitle("Multiverse of All Reasonable Meta-Analyses") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_text(colour = c("white", rep("black", times = length(y_labels_forest) - 1))),
        axis.ticks.y = element_line(colour = c("white", rep("black", times = length(y_breaks_forest) - 1))),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(),
        panel.grid.minor.y = element_blank(),
        plot.margin = margin(t = 5.5, r = 5.5, b = -15, l = 5.5, unit = "pt")) +
  geom_label_repel(data = ma_for_forest, 
                   aes(label = label),
                   seed = 42,
                   box.padding   = 0.35, 
                   point.padding = 0.5,
                   direction = "both",
                   lineheight = 2,
                   nudge_y = .5,
                   segment.color = 'black',
                   segment.angle = 90,
                   segment.size = .5)

p2_rel
```

#### Combine specfication tile plot, subset size indicator and forest plot 

```{r}
skelleton <- plot_grid(effect_pane,
                   factor_pane,
                   ncol = 1,
                   align = "v",
                   rel_heights = c(4,5))
skelleton

ggsave("figures/descriptive_skelleton.png", 
       skelleton, 
       width = 10, 
       height = 7.5,
       dpi = "retina"
       )
```

```{r}
p_rel <- plot_grid(p2_rel,
                   p1_rel,
                   ncol = 1,
                   align = "v",
                   rel_heights = c(4,5))
p_rel
```

<br>

## Save Plot

```{r }
ggsave("figures/descriptive_specification.png", 
       p_rel, 
       width = 10, 
       height = 7.5,
       dpi = "retina"
       )
```
